### 2015-02-06 cav-2015-tex/rw.tex
Did another pass on the related work.

To address the issue of references, several projects have generalized
the core idea of symbolic execution, enabling it to be applied to
programs with more general types, including references and
arrays~\cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. However, these generalized symbolic execution (GSE) techniques work by branching over multiple copies of the system state during dereferencing operations, thus exacerbating the path explosion problem. Our work leverages the core idea in generalized symbolic execution
with lazy initialization to construct the heap on-the-fly during
symbolic execution, but avoids the nondeterminism introduced by GSE by
constructing a summary heap and will generally create fewer paths than
GSE techniques.

@@

Dynamic symbolic execution (\dsetxt{}, sometimes referred to as concolic execution) has emerged as a way to enable symbolic execution to continue on programs that contain elements that cannot be reasoned about symbolically. ~\dsetxt{} works by executing a program using concrete and symbolic engines simultaneously. When the program reaches a statement that cannot be reasoned about symbolically, it substitutes in values from the concrete execution and continues. When reaching the end of a program path, DSE inverts the branch conditions and uses the solver to generate concrete inputs for another execution path. While concretization allows DSE to explore a much wider selection of programs, this freedom comes at the price of completeness. The process of concretization introduces approximation such that all possible execution paths may not be explored.

In the absence of an effective theory of references, DSE has been a natural choice for reasoning about programs that contain them. A number of DSE techniques use concretization when reasoning about references~\cite{Godefroid:2005,Sen:2005,Godefroid:POPL07,Tillmann:2008}. However, as demonstrated in~\cite{Elkarablieh:2009}, DSE underapproximates when it concretizes symbolic values, resulting in an incomplete analysis. The SAGE DSE engine~\cite{Elkarablieh:2009} includes more sophisticated methods for reasoning about references, and can successfully reason about more complicated examples. However, the heap model in SAGE requires fixing the size of memory prior to starting the analysis, and is therefore not complete.

\noindent\textbf{Bounded Model Checking}
Bounded model checking (BMC) techniques seek to avoid the path explosion problem by converting an entire program execution, along with its specifications, into a constraint problem called a verification condition. The advantage in creating a single constraint is that SMT solvers can often reason more efficiently about all paths simultaneously rather than on a case-by-case basis. Many BMC methods include a model for reasoning about references, including those described in~\cite{Clarke:2004,Barnett:2006,Xie:2005,Babic:2007,Dillig:2011}. However, BMC techniques are not complete for general programs, due to the need to set an execution path-length bound prior to conducting the analysis. By extension, the bounded reference models used in BMC methods are generally not complete. Other limitations include assuming that symbolic references do not alias~\cite{Xie:2005,Babic:2007}, that data structures are not recursive~\cite{Dillig:2011}, or requiring that symbolic references are never dereferenced~\cite{Clarke:2004}.

\noindent\textbf{Other Symbolic Heap Representations}
%Due to technological limitations, many early symbolic heap representations focused on reducing the computational complexity of dereferencing queries. By introducing certain heuristics and approximations, the computational burden can be greatly reduced, at the cost of precision. Since the problem of weeding out false positives can be at least as difficult as the program analysis itself, there has been a renewed interest in more precise heap representations. However, these newer methods generally err on the side of under-approximation, producing an analysis that is not complete.
%
%Several GSE techniques~\cite{GSE03,KiasanKunit} form an exception to this rule, presenting an input heap model that is unconstrained and unbounded. However, GSE does not offer a mechanism for representing the set of heaps along an execution path as a unified symbolic value. As a result GSE does not produce a path condition representing the complete set of inputs for a given execution path, negating a significant advantage of symbolic execution. The work presented here overcomes this limitation by using symbolic values to represent the heap.
A number of constraint-based programming techniques are capable of reasoning about heaps~\cite{Degrave:2010,Charreteur:2009,Albert:2013}. Most notably, the technique in~\cite{Albert:2013} is capable of reasoning about dereferencing operations in a per-path set of unknown input heaps. However, none of these methods provide guarantees about the precision of the analysis performed, nor is it clear whether such guarantees are possible.

% a good place to mention procedure summaries, like Yorsh and the Dilligs.
The symbolic heap value sets presented in this work build upon a number of prior techniques that use sets of guarded values to represent program state~\cite{Sen:2014,Torlak:2014,Yorsh:2008,Xie:2005,Dillig:2011,Elkarablieh:2009}. The methods presented in~\cite{Sen:2014,Torlak:2014,Yorsh:2008} use value sets to represent all symbolic program state, including references. While symbolic referencing operations could, in principle, be supported with these methods, it is not clear from the published material exactly how that might be accomplished, especially in the case of dereferencing. All of the methods presenting solutions to the dereferencing problem~\cite{Xie:2005,Dillig:2011,Elkarablieh:2009} are not complete, for a variety of reasons which have been enumerated previously.

>>>>

To address the issue of references, several projects have generalized
the core idea of symbolic execution, enabling it to be applied to
programs with more general types, including references and
arrays~\cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. However, these generalized symbolic execution (GSE) techniques work by branching over multiple copies of the system state during dereferencing operations, exacerbating path explosion and fragmenting the path condition. Producing a complete path condition for a single given control flow path would require merging an exponential number of states. Our work leverages the core idea in generalized symbolic execution with lazy initialization using on-the-fly reasoning to model a black-box input heap during
symbolic execution, but avoids the nondeterminism introduced by GSE by
constructing a single summary heap and polynomially-sized path condition for each control flow path.

@@

Dynamic symbolic execution (\dsetxt{}) has emerged as a way to enable symbolic execution to continue on programs that contain elements that cannot be reasoned about symbolically. ~\dsetxt{} works by executing a program using concrete and symbolic engines simultaneously. When the program reaches a statement that cannot be reasoned about symbolically, it substitutes in values from the concrete execution and continues. In the absence of an effective theory of references, DSE has been a natural choice for reasoning about programs that contain them. A number of DSE techniques use concretization when reasoning about references~\cite{Godefroid:2005,Sen:2005,Godefroid:POPL07,Tillmann:2008}. However, as demonstrated in~\cite{Elkarablieh:2009},~\dsetxt{} underapproximates when it concretizes symbolic values, resulting in an incomplete analysis. The SAGE~\dsetxt{} engine~\cite{Elkarablieh:2009} includes more sophisticated methods for reasoning about references, and can successfully reason about more complicated examples. However, the heap model in SAGE prohibits input-dependent memory allocation. The fixed-memory requirement means that analyses of programs accepting indefinite input heaps may terminate prematurely, even for the class of self-limiting solver-compatible programs. Since symbolic execution is otherwise sound and complete for this class, bounded heap models are, in general, incomplete with respect to symbolic execution. In contrast, the~\symtxt{} model presented in this paper represents a truly unbounded black box. While practical considerations may require that some sort of bounds be placed on the symbolic execution itself, the nature of those bounds need not be dictated by the~\symtxt{}.

\noindent\textbf{Bounded Model Checking}
Bounded model checking (BMC) techniques seek to avoid the path explosion problem by converting an entire program execution, along with its specifications, into a constraint problem called a verification condition. The advantage in creating a single constraint is that SMT solvers can often reason more efficiently about all paths simultaneously rather than on a case-by-case basis. Many BMC methods include a model for reasoning about references, including those described in~\cite{Clarke:2004,Barnett:2006,Xie:2005,Babic:2007,Dillig:2011}. However, BMC techniques are not complete for general programs, due to the need to set an execution path-length bound prior to conducting the analysis. By extension, the reference models used in BMC methods generally require a predetermined size bound, and and are thus incomplete with respect to symbolic execution. Other limitations include assuming that symbolic references do not alias~\cite{Xie:2005,Babic:2007}, that data structures are not recursive~\cite{Dillig:2011}, or prohibiting dereferencing of black-box input references~\cite{Clarke:2004}.

\noindent\textbf{Other Symbolic Heap Representations}
%Due to technological limitations, many early symbolic heap representations focused on reducing the computational complexity of dereferencing queries. By introducing certain heuristics and approximations, the computational burden can be greatly reduced, at the cost of precision. Since the problem of weeding out false positives can be at least as difficult as the program analysis itself, there has been a renewed interest in more precise heap representations. However, these newer methods generally err on the side of under-approximation, producing an analysis that is not complete.
%
%Several GSE techniques~\cite{GSE03,KiasanKunit} form an exception to this rule, presenting an input heap model that is unconstrained and unbounded. However, GSE does not offer a mechanism for representing the set of heaps along an execution path as a unified symbolic value. As a result GSE does not produce a path condition representing the complete set of inputs for a given execution path, negating a significant advantage of symbolic execution. The work presented here overcomes this limitation by using symbolic values to represent the heap.
A number of constraint-based programming techniques are capable of reasoning about heaps~\cite{Degrave:2010,Charreteur:2009,Albert:2013}. Most notably, the technique in~\cite{Albert:2013} is capable of reasoning about dereferencing operations in a per-path set of unknown input heaps. However, no guarantee is made about the precision of the analysis performed, nor is it obvious that such a guarantee is possible for the proposed method.

% a good place to mention procedure summaries, like Yorsh and the Dilligs.
The symbolic heap value sets presented in this work build upon a number of prior techniques that use sets of guarded values to represent program state~\cite{Sen:2014,Torlak:2014,Yorsh:2008,Xie:2005,Dillig:2011,Elkarablieh:2009}. The methods presented in~\cite{Sen:2014,Torlak:2014,Yorsh:2008} use value sets to represent all symbolic program state, including references. While symbolic referencing operations could, in principle, be supported with these methods, it is not clear from the published material exactly how that might be accomplished, especially when reasoning about dereferencing in a black-box input heap. All of the methods presenting solutions to the dereferencing problem~\cite{Xie:2005,Dillig:2011,Elkarablieh:2009} are not complete, for a variety of reasons which have been previously enumerated.
